Exploring probability distributions for bivariate temporal granularities

### 2019-11-20 main.Rmd
cricket application update

## Single-order-up granularities

@@

## Multiple order-up granularities

@@

# Case studies

@@

The ball by ball data for IPL season 2008 to 2016 from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).  The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).

@@

Q1: How run rates vary depending on if a team bats first or second?
Q2: How number of wickets and dot balls in the previous over affect the runs in the subsequent over across different overs of the innings?

@@

A dot ball is a delivery bowled without any runs scored off it. The number of dot balls is reflective of the quality of bowling in the game. Run rate of an over should ideally decrease if the number of dot balls increase or the batsman can likely to go for big shots and get dismissed. Figure \ref{fig:dot} shows that for any over, increase in dot balls lead to decrease in run rate.

@@

```{r exdotplot, echo=FALSE}
library(gravitas)
library(tsibble)

cricket_bowling <- cricket_season %>% 
  mutate(dot_ball_proxy = if_else(total_runs==0,1,0)) %>% 
  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
  summarise(runs_per_over = sum(total_runs),
            run_rate = sum(total_runs)/length(total_runs),
            dot_balls = sum(dot_ball_proxy)) %>%
            mutate(lagged_run_rate = c(0,diff(run_rate)))

@@

cricket_tsibble_bow %>% 
  filter(dot_balls %in% c(0, 1, 2))%>% 
  prob_plot("dot_balls",
            "over",
            hierarchy_model,
            response = "lagged_run_rate",
            plot_type = "quantile",
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

@@

```{r exwicketplot}
cricket_tsibble %>% 
  filter(wicket %in% c(0, 1)) %>%
  mutate(lag_wicket = lag(wicket)) %>% 
  filter(lag_wicket %in% c(0, 1)) %>%
  prob_plot("lag_wicket",
            "over",
            hierarchy_model,
            response = "runs_per_over",
            plot_type = "quantile",
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

@@

{r read_data, echo=FALSE}

>>>>

### Single-order-up granularities

@@

### Multiple order-up granularities

@@

# Applications

@@

The ball by ball data for IPL season 2008 to 2016 is fetched from [Kaggle](https://www.kaggle.com/josephgpinto/ipl-data-analysis/data).  The `cricket` data set in the `gravitas` package summarizes the ball-by-ball data cross overs and contains information for a sample of 214 matches spanning 9 seasons (2008 to 2016).

@@

Q1: How run rates vary depending on if a team bats first or second?

Q2: How number of wickets and dot balls in the previous over affect the runs in the subsequent over across different overs of the innings?

@@

A dot ball is a delivery bowled without any runs scored off it. The number of dot balls is reflective of the quality of bowling in the game. Run rate of an over should ideally decrease if the number of dot balls increase. However, what is the effect of dot balls in previous over on run rate in the subsequent over. Will players  batsman likely to go for big shots because they couldn't score good runs in the previous over? Or they should play consistently and avoid scoring high?

Figure \ref{fig:dot} shows that for any over, increase in dot balls lead to decrease in run rate.

@@

```{r exdotplot, echo=FALSE, eval = FALSE}
library(gravitas)
library(tsibble)

cricket_bowling <- cricket_season %>% 
  mutate(dot_ball_proxy = if_else(total_runs==0,1,0)) %>% 
  group_by(season, match_id, batting_team, bowling_team,  inning, over) %>%
  summarise(runs_per_over = sum(total_runs),
            run_rate = sum(total_runs)/length(total_runs),
            dot_balls = sum(dot_ball_proxy)) %>%
            mutate(diff_run_rate = c(0,diff(run_rate)))

@@

cricket_tsibble_bow %>% 
  filter(dot_balls %in% c(0, 1, 2))%>% 
  prob_plot("dot_balls",
            "over",
            hierarchy_model,
            response = "diff_run_rate",
            plot_type = "quantile",
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

@@

```{r exwicketplot}

cricket_tsibble %>% 
  filter(dot_balls %in% c(0, 1, 2)) %>%
  mutate(lag_dot = lag(dot_balls)) %>% 
  filter(lag_dot %in% c(0, 1, 2)) %>%
  prob_plot("lag_dot",
            "over",
            hierarchy_model,
            response = "runs_per_over",
            plot_type = "quantile",
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

cricket_tsibble %>% 
  filter(wicket %in% c(0, 1)) %>%
  mutate(lag_wicket = lag(wicket)) %>% 
  filter(lag_wicket %in% c(0, 1)) %>%
  prob_plot("lag_wicket",
            "over",
            hierarchy_model,
            response = "runs_per_over",
            plot_type = "quantile",
            quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))

@@

<!-- {r read_data, echo=FALSE} -->
